\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-03-16 write_up/mixfishsim_ecolmod.tex
Updates to manuscript and developments on results

@@

We simulate 20 years of exploitation of the fish populations and use the
results from the fishing model to draw inference on the underlying population
structures. We compare this inference to i) a simulated fixed-site sampling
design commonly used for fisheries monitoring purposes, and ii) the true
underlying population structures input to the simulation, to establish the
potential and limitations of fishery-dependent data - an inherently biased
sampling method due to fisher's targeting- to provide a robust picture of
spatiotemporal distributions. [We simulate a closure based on areas defined
from commercial catch data and assess its effectiveness on reducing catches of
a fish population.] \\

We conclude that... \\

@@

[Could fit a geostatistical model (e.g. VAST) to the fisheries-dependent and
fisheries-independent data, though may be overkill...] \\

@@

We parameterised the simulation model for four populations with differing
habitat preference (Figure \ref{fig:1}), population demographic and recruitment
functions; each of the populations also has two defined spawning areas and
movement rates (Table \ref{tab:1}). \\

@@

The fleets were parameterised to reflect five different characteristics based
on targeting preference and exploitation dynamics (Table \ref{tab:2}). This
ensures that different fleets have different spatial dynamics, preferentially
targeted different fish populations. The stochasticity in the random walk
process ensures that different vessels within a fleet have slightly different
spatial distributions based on individual experience, while the step function
was parameterised dynamically so that vessels take smaller steps where the
fishing location yields in the top X \textcolor{red}{[??]}quartile of the value
available in that year (as defined per fleet in Table \ref{tab:2}). \\

@@

@@

The survey simulation was set up with follow a fixed gridded station design
with 49 stations fished each year, starting on day 92 with same catchability
parameters for all populations (Q = 1).

@@

We allow the simulation to run unrestricted for 5 years, and subsequently close
areas for the last 5 years of the simulation based on data (either derived from
the commercial catches, fisheries-independent survey or the 'real population')
used at different spatial and temporal scales. \\

@@

In total 56 closure scenarios were run which represent combinations of: three
data types; commercial logbook data, survey data and 'real population', three
temporal resolutions; weekly, monthly and yearly closures, and four spatial
resolutions; 1 x 1 grid, 5 x 5 grid, 10 x 10 grid and 20 x 20 grid. Survey
closures were on an annual basis, as the most temporally resolved survey data
available.

@@

Need to consider what best to present here as 4 / 5 figures:

\begin{itemize}
	\item Spatial dynamics: e.g. Figure 14. showing the population movement
		across weeks, including a spawning period.
	\item Overall population trends: e.g. Figure 3, showing the population
		dynamics at play.
	\item Realised step function: e.g. Figure 13, showing how the movement
		and turning angle responds. This could be combined with Figure
		7 for a more complete visualisaiton.
	\item Catch composition and spatial clustering: e.g. Figure 10, showing
		the importance of spatial resolution.
	\item Some measure of temporal trends/changes in spatial composition -
		could pick a certain cell and show how the proportions change
		over time ? An example is Figure 15, though there is little
		variation here. I think this is because the static habitat with
		population movement is far more stable than the previous
		directly affected population distributions (where there was an
		advective-diffusive process from the SPDE). Might need to
		consider how to change this?  Could include a spatiotemporally
		changing habitat suitability covariate, e.g. temperature ??
		\textcolor{red}{[temperature would be very interesting and
			cover varying spatial fields]}
	\item Comparison of population structure from i) commercial sampling,
		ii) fisheries independent survey, iii) real population. This
		should be some statistical measure...not sure best approach
		here.
	\item [If we perform the closure, some measure of cost/benefit based on
		different resolutions of data and data used (i.e. commercial,
		survey, real populations). This could be on the basis of
		closing x cells, giving cell references to the model so they
		cannot be fished. Need to consider whether this is also a
		seasonal closure.]
\end{itemize}

Present simulated closures in terms of \% change in population biomass and
fishery.

@@

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/Data_Aggregation_space}
	\caption{Data aggregation at different spatial resolutions}
	\label{fig:0a}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/Data_Aggregation_time}
	\caption{Data aggregation at different temporal resolutions}
	\label{fig:0b}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/Overview_plot_highPop}
	\caption{Comparison of closure scenarios}
	\label{fig:0c}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/F_trends}
	\caption{Comparison of closure scenarios - F trends}
	\label{fig:0d}
\end{figure}

// 
//Supplementary
//

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/habitat}
	\caption{habitat preference}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/annual_summary}
	\caption{Summary of annualised metrics: biomass, catch, fishing
		mortality and recruitment. x-axis is the year, y the value}
	\label{fig:2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/fDynamics}
	\caption{f dynamics - the daily fishing mortalities, each year is a
		different colour}
	\label{fig:3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/vessel_move}
	\caption{vessel movement - a single trip movement for one vessel}
	\label{fig:4}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/vessel_multi_move}
	\caption{vessel movement for multiple trips from a single vessel. Note
	the movement off the side pops up the other side, but is joined by a
	line across the grid. This is from the torus approach rather than the
	edges being barriers}
	\label{fig:5}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/vessel_move_value}
	\caption{movement of a single vessel over a few trips overlaid on the
		value field (i.e. sum of the population abundance x catchability x value}
	\label{fig:6}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/fleet_moves}
	\caption{An entire fleets (20 vessels) movement for a single trip}
	\label{fig:7}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/catch_comp}
	\caption{spatial catch composition - the raw catches per cell at 4
		different spatial resolutions. Sorry, its a bit small at the
		moment!}
	\label{fig:8}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/catch_comp_clusters}
	\caption{spatial catch composition (as above) but with clustering of
		cells performs on the bottom row. I'm confused as to why some
		entirely blue cells in the raw catch composition get allocated
	to the same cluster as some entirely red cells - needs investigating,
	it may be the clustering is not optimal}
	\label{fig:9}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/survey_index}
	\caption{survey index - a non-spatial index generated from the fishery independent
	survey}
	\label{fig:10}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/spatial_survey}
	\caption{survey spatial abundance - a bubble plot of the survey
		abundances, not really useful for the paper}
	\label{fig:11}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/step_function}
	\caption{Realised step function - the step function as realised for a
		single fleet. For turning angles, it can be seen that at higher
	values, the turning range is less}
	\label{fig:12}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/pop_dist}
	\caption{Spatial dynamics - the four populations at four time steps.
		Not really happy about the lack of dynamics here - mainly down
	to the static habitat leading to the populations settle in same cells
	over most time steps (exception: during the spawning period). Could
	include a temperature covariate or something to force some more
	temporal changes?}
	\label{fig:13}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/Proportion_in_cell}
	\caption{Temporal dynamics - the proportion of each population in a
		randomly chosen cell. Similar criticism as above}
	\label{fig:14}
\end{figure}

>>>>

\usepackage{xr} %% For cross-referencing supl figures
\externaldocument[S-]{Supplementary}

@@

We simulate 20 years of exploitation of the fish populations and use the
results from the fishing model to draw inference on the underlying population
structures. We compare this inference to i) a simulated fixed-site sampling
design commonly used for fisheries monitoring purposes, and ii) the true
underlying population structures input to the simulation, to establish the
potential and limitations of fishery-dependent data - an inherently biased
sampling method due to fisher's targeting- to provide a robust picture of
spatiotemporal distributions. Finally, we simulate an area closure based on
areas defined from commercial the known ("real-population") distribution,
commercial catch data and survey data at different temporal and spatial
resolutions and assess their effectiveness on reducing catches of a fish
population. \\

We conclude from our simulations that commercial data, while not unbiased,
provides a useful tool for managing catches in mixed fisheries if applied at
the correct spatiotemporal scale. \\

@@

@@

We parameterised the simulation model for four populations with differing
habitat preference and temperature tolerances (Figures \ref{S-fig:1},
\ref{S-fig:3}, \ref{S-fig:4}, \ref{S-fig:5}, \ref{S-fig:6}, \ref{S-fig:7}),
population demographic and recruitment functions; each of the populations also
has two defined spawning areas (Figure \ref{S-fig:2}) and movement rates (Table
\ref{tab:1}). The actual movement of the populations for a number of weeks is
shown in Figure \ref{S-fig:9} while the realised daily fishing mortality are
shown in Figure \ref{S-fig:10}. \\

@@

The fleets were parameterised to reflect five different characteristics based
on targeting preference and exploitation dynamics (Table \ref{tab:2}). This
ensures that different fleets have different spatial dynamics, preferentially
targeted different fish populations. The stochasticity in the random walk
process ensures that different vessels within a fleet have slightly different
spatial distributions based on individual experience, while the step function
was parameterised dynamically so that vessels take smaller steps where the
fishing location yields in a top quartile of the value available in that year
(as defined per fleet in Table \ref{tab:2}). \\

@@

An example of the realised fleet movements for a single vessel during a single
trip are given in Figure \ref{S-fig:11}, while Figure \ref{S-fig:12} shows
multiple trips for a single vessel, \ref{S-fig:13} the vessel movements for
some trips overlaid on the value field, \ref{S-fig:14} shows fishing locations
for an entire fleet of 20 vessels for a single trip, while \ref{S-fig:15} shows
an example of the step function realisation and turning angles from the
correlated random walk.

@@

The survey simulation was set up with follow a fixed gridded station design
with 100 stations fished each year, starting on day 92 with same catchability
parameters for all populations (Q = 1).

@@

We allow the simulation to run unrestricted for 5 years, and subsequently close
areas for the last 5 years of the simulation based on data (either derived from
the commercial catches, fisheries-independent survey or the 'real population' -
the underlying populations assumed to be known perfectly) used at different
spatial and temporal scales. \\

@@

In total 56 closure scenarios were run which represent combinations of

\begin{itemize}
	\item \textbf{data types:} commercial logbook data, survey data and
		'real population',
	\item \textbf{temporal resolutions:} weekly, monthly and yearly
		closures,
	\item \textbf{spatial resolutions:} 1 x 1 grid, 5 x 5 grid, 10 x 10
		grid and 20 x 20 grid.
\end{itemize}

Survey closures were on an annual basis only, as this was the most temporally
resolved survey data available.

@@

The species distribution themselves

The consequences of different spatial aggregations of the data are shown in
Figure \ref{fig:1}, which represents the aggregation of catch from each of the
data sources over a year at different spatial resolutions. \\

The finer spatial grid for the the real population (top left) and commercial
data (top middle) show similar patterns, though there are unsampled gaps in the
commercial data from a lack of fishing activity (particularly in the lower left
part of the sampling domain). The survey data at this spatial resolution shows
very sparse and uninformative information about the spatial distributions of
the populations. The slightly aggregated data on a 5 x 5 grid shows similar
patterns, and while losing some of the spatial detail there remains good
consistency between the 'real population' and the commercial data. Survey data
starts to pick out some of the similar patterns as the other data sources, but
lacks coverage. The spatial catch information on a 10 x 10 and 20 x 20 grid
loses a significant amount of information about the spatial resolutions for all
data sources, and some differences between the commercial and 'real population'
data emerge. \\

Figure \ref{fig:2} shows the consequences of different temporal aggregations of
the data, with 156 weekly (top), 36 monthly (middle) and 3 yearly (bottom)
catch compositions across a 20 x 20 area. \\

As can be seen from the 'real population', the monthly aggregation captures the
major patterns seen in the weekly data, albeit missing more subtle differences.
The yearly data results in a constant catch pattern due to the aggregation
process (sometimes known as an aggregation bias). The commercial data on a
weekly basis shows some of the same patterns as the 'real population', though
the first species (in red) is less well represented and some weeks are missing
catches from the area. The monthly data. The monthly data shows some
consistency between the 'real population' and commercial data for species 2 -
4, though species 1 remains underrepresented. On an annual basis, interestingly
the commercial data underrepresents the first species (in red) while the survey
overrepresents species 1. This is likely due to the biases in commercial
sampling, with the fisheries not targeting the areas where species 1 are
present, and the biases in the survey sampling from overrepresentation of the
spatial distribution. \\

We implemented a spatial closure using the different data sources and spatial
and temporal aggregations as outlined in the protocol in Section 3.4. We used
this to assess the efficacy of a closure in reducing fishing mortality on
species 1, given availability of data and its use at different resolutions in
order to evaluate the trade-offs in data sources. Figure \ref{fig:3} shows the
trend in fishing mortality for each species simulated (columns) given the data
sources (rows), temporal aggregations (colour lines) and spatial aggregations
(linestyles), while Figure \ref{fig:4} shows the change in fishing mortality
from before the closure (average F years 2 - 4) to after the closure (average F
years 8 - 10). \\

For the closures based on 'real population' (bottom row), the most
disaggregated data (a weekly timescale and 1 x 1 resolution) was most
effective, reducing fishing mortality on species 1 (left) by $\sim$ 60 \%. Next
was the monthly closures ($<$ $\sim$ 30 \%).  The least effective were the
yearly closures (blue lines) at all spatial resolutions, which resulted in
increased fishing mortalities ($>$ 30 \% - \textcolor{blue}{N.B. Note though,
this is consistent with the increasing trends in F, which is probably more
related to the fact that Fs hadn't stabilised in the simulation from the
fishing vessels "learning" the best locations - I will rerun the sims for a
longer time (20 - 30 years)}. \\

For the survey data, which can only be implemented on a yearly timescale, the
closures had no effect at any data resolution. The results are identical for
the different data resolutions except 20 x 20, which is why you can't see more
than 2 points. This is because of the sparsity of the sampling locations.\\

For the commercial data, the most effective closure scenario was based on 1 x 1
data at a monthly temporal resolution. This results in $\sim$ 10 \% reduction
in F for species 1. \textcolor{blue}{This was the only closure scenario to have
	positive effect according to Figure \ref{fig:4}, though looking at the
	trend in Figure \ref{fig:3} this looks more related to the continued
	increased in F trend, as other scenarios had an initial effect}.
Interestingly the monthly data scenario was more effective than weekly data,
which I'd posit is due to the increase amount of data available from the
commercial sampling across a month compared to a week.i Commercial data used at
an annual timestep was ineffective in bringing fishing mortality down for
species 1. \\

Given the scenarios above, it seems clear that spatial disaggregation is more
important than the temporal disaggregation of the commercial data, except when
its used at an annual timeframe, which is the scenario that gave the worst
results.

For the other species in the simulation (population 2 - 4) there was little
difference in fishing mortalities across scenarios.

\textcolor{blue}{Note: The monthly commercial data scenario is the most
	effective of the realistic scenarios, as the 'real population' can only
	be seen as a baseline comparison.}

@@

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/Data_Aggregation_space}
	\caption{Data aggregation at different spatial resolutions}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/Data_Aggregation_time}
	\caption{Data aggregation at different temporal resolutions}
	\label{fig:2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/F_trends}
	\caption{Comparison of closure scenarios - F trends}
	\label{fig:3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/Overview_plot_highPop}
	\caption{Comparison of closure scenarios}
	\label{fig:4}
\end{figure}
